
---
title: "Muscle Mass Regression Analysis"
output:
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: scroll
---

```{r setup, include=FALSE}
library(ggplot2)
library(plotly)
library(dplyr)
library(flexdashboard)
library(readr)
library(knitr)
library(DT)
library(shiny)

# Load the dataset
CH01PR27 <- read_table2("CH01PR27.txt", col_names = c("Y", "X"))

# Center the variable X for higher-order models
CH01PR27 <- CH01PR27 %>%
  mutate(x_centered = X - mean(X))
```

Row {data-orientation=rows}
-----------------------------------------------------------------------

### Interactive Data Exploration

A person's muscle mass is expected to decrease with age. To explore this relationship in women, a nutritionist randomly selected 15 women from each 10-year age group, beginning with age 40 and ending with age 79. The results follow; $X$ is age, and $Y$ is a measure of muscle mass. Assume that first-order regression model (1.1) is appropriate.
\begin{tabular}{rrrrrrrr}
$i:$ & \multicolumn{1}{c}{$\mathbf{1}$} & \multicolumn{1}{c}{$\mathbf{2}$} & $\mathbf{3}$ & $\ldots$ & $\mathbf{5 8}$ & $\mathbf{5 9}$ & $\mathbf{6 0}$ \\
\hline$X_i:$ & 43 & 41 & 47 & $\ldots$ & 76 & 72 & 76 \\
$Y_i:$ & 106 & 106 & 97 & $\ldots$ & 56 & 70 & 74
\end{tabular}

**Explore the dataset interactively:**

```{r}
# Create an interactive data exploration widget
datatable(
  CH01PR27,
  options = list(pageLength = 10, autoWidth = TRUE),
  caption = "Interactive Table of Muscle Mass Data"
)
```

**Customize the Scatter Plot:**

```{r}
# Add an input widget for group selection
selectInput(
  inputId = "group_selection",
  label = "Choose Age Range:",
  choices = c("40-49", "50-59", "60-69", "70-79"),
  selected = "40-49"
)

# Filter dataset based on age range
filtered_data <- reactive({
  age_ranges <- list(
    "40-49" = CH01PR27 %>% filter(X >= 40 & X < 50),
    "50-59" = CH01PR27 %>% filter(X >= 50 & X < 60),
    "60-69" = CH01PR27 %>% filter(X >= 60 & X < 70),
    "70-79" = CH01PR27 %>% filter(X >= 70 & X <= 79)
  )
  age_ranges[[input$group_selection]]
})

# Interactive scatter plot for the selected age range
renderPlotly({
  ggplot(filtered_data(), aes(x = X, y = Y)) +
    geom_point(color = "blue", alpha = 0.7) +
    labs(
      title = paste("Muscle Mass vs. Age (", input$group_selection, ")", sep = ""),
      x = "Age (X)", y = "Muscle Mass (Y)"
    ) +
    theme_minimal()
})
```

Row {data-orientation=rows}
-----------------------------------------------------------------------

### First-Order Regression

**Problem Statement**:  
A nutritionist wants to explore the relationship between muscle mass (\(Y\)) and age (\(X\)) in women. The goal is to fit a first-order regression model and analyze the relationship.

#### Solution and Visualization

```{r}
#  Problem 1.27:  Fit the first-order regression model
model1 <- lm(Y ~ X, data = CH01PR27)
summary1 <- summary(model1)

# Display regression summary
kable(summary1$coefficients, caption = "Summary of First-Order Regression Coefficients")

# Plot scatter and regression line
p1 <- ggplot(CH01PR27, aes(x = X, y = Y)) +
  geom_point(color = "blue", alpha = 0.7) +
  geom_smooth(method = "lm", color = "red", se = FALSE) +
  labs(
    title = "First-Order Regression: Muscle Mass vs. Age",
    x = "Age (X)", y = "Muscle Mass (Y)"
  ) +
  theme_minimal()

# Convert to interactive plot
ggplotly(p1)
```


### Second-Order Regression

**Problem Statement**:  
Extend the analysis by fitting a second-order (quadratic) regression model and evaluating its fit. Additionally, predict the mean and individual muscle mass for a 48-year-old woman.

#### Solution and Visualization

```{r}
# Problem 8.4: Fit second-order regression model
model2 <- lm(Y ~ x_centered + I(x_centered^2), data = CH01PR27)
summary2 <- summary(model2)

# Display regression summary
kable(summary2$coefficients, caption = "Summary of Second-Order Regression Coefficients")

# Generate predictions for visualization
X.new <- seq(min(CH01PR27$x_centered), max(CH01PR27$x_centered), length.out = 100)
fit2 <- predict(model2, newdata = data.frame(x_centered = X.new))
fit_data <- data.frame(x_centered = X.new, fit2)

# Plot data with quadratic regression line
p2 <- ggplot(CH01PR27, aes(x = x_centered, y = Y)) +
  geom_point(color = "green", alpha = 0.7) +
  geom_line(data = fit_data, aes(x = x_centered, y = fit2), color = "blue") +
  labs(
    title = "Second-Order Regression: Quadratic Fit",
    x = "Centered Age (X)", y = "Muscle Mass (Y)"
  ) +
  theme_minimal()

# Convert to interactive plot
ggplotly(p2)

# Confidence and prediction intervals for age 48
age_centered <- 48 - mean(CH01PR27$X)
confidence_interval <- predict(model2, newdata = data.frame(x_centered = age_centered), interval = "confidence")
prediction_interval <- predict(model2, newdata = data.frame(x_centered = age_centered), interval = "prediction")

# Display intervals as a table
interval_data <- data.frame(
  Type = c("Confidence Interval", "Prediction Interval"),
  Lower = c(confidence_interval[2], prediction_interval[2]),
  Upper = c(confidence_interval[3], prediction_interval[3])
)
kable(interval_data, caption = "95% Confidence and Prediction Intervals for Age 48")
```

Row {data-orientation=rows}
-----------------------------------------------------------------------

### Residual Diagnostics: Normal Q-Q Plot


#### Visualizing Normality of Residuals:
```{r}
# Fit the second-order regression model
model2 <- lm(Y ~ x_centered + I(x_centered^2), data = CH01PR27)

# Normal Q-Q plot
residuals <- resid(model2)
p_qq <- ggplot(data.frame(Residuals = residuals), aes(sample = Residuals)) +
  stat_qq() +
  stat_qq_line(color = "blue") +
  labs(
    title = "Normal Q-Q Plot for Residuals",
    x = "Theoretical Quantiles",
    y = "Sample Quantiles"
  ) +
  theme_minimal()

# Convert to interactive plot
ggplotly(p_qq)

```

### Residuals vs. Fitted Values:


```{r}
# Residuals vs. fitted values plot
fitted_values <- fitted(model2)
p_residuals <- ggplot(data.frame(Fitted = fitted_values, Residuals = residuals), aes(x = Fitted, y = Residuals)) +
  geom_point(color = "purple", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
  labs(
    title = "Residuals vs. Fitted Values",
    x = "Fitted Values",
    y = "Residuals"
  ) +
  theme_minimal()

# Convert to interactive plot
ggplotly(p_residuals)
```


